Visualizing probability distributions across bivariate cyclic temporal granularities

### 2021-05-13 main.Rmd
captions for last three figures re-arranged

(ref:allFig) Distribution of energy consumption displayed through letter value plots to illustrate harmonies and clashes and how change in x-axis and facet mappings change emphasis. **c** shown across month-of-year faceted by quarter-of-year is a clash, as there are empty sets. For example, November can never correspond to Quarter 1. **a** shown across quarter-of-year faceted by weekday/weekend do not include any empty sets and hence is a harmony. The same is true for **b**, which shows the same two cyclic granularities as **a** with reverse mapping. There are no empty combination since each quarter contains both weekdays and weekends. In **a**, it is easier to compare across weekday type (x-axis) within a quarter (facet); within quarter 2, more mass occupied the lower tail on the weekends (letter value E with depth 1/8) relative to that of the weekdays (letter value D 1/16). Coversely, **b** exaggerates differences across quarters (x-axis now) within a specific weekday type; within weekends, the lower letter values C and D for the first and forth quarter are higher than the other quarters.
This is probably because of temperatures in those two quarters are higher as they correspond to summer and spring in Australia.

@@

(ref:bothcust) Energy consumption of a single customer shown with different distribution displays, and granularity arrangements. Two granularities are used: 1. hour of the day; and 2. weekday/weekend. **a** shows granularity 2 faceted by granularity 1; **b** and **c** shows the reverse mapping. **a** allows comparison of usage by workday within each hour of the day using side-by-side boxplots, showing that on work days there is more consumption early in the day. **b** and **c** examine the temporal trend of consumption over the course of a day, separately for the type of day. **b** uses an area quantile to emphasize the time component; e.g., median consumption shows prolonged high usage in the morning on weekdays. **c** uses a violin plot emphasizes subtler distributional differences across hours, showing that morning use on weekdays is bimodal, and on some work days there is low usage possibly indicating the person is working from home or having a late start.

@@

(ref:cricex) Distributions of runs per over shown with different distribution displays, and granularities. **a** shows letter value plots across overs faceted by innings. For the team batting in the first innings, there is an upward trend of runs per over, while there is no such pattern of runs for the teams batting in the second innings. **b** shows quantile plots of runs per over across an indicator of wickets in the previous over faceted by current over. This indicates that at least one wicket in the previous over leads to a lower median run rate and quartile spread in the subsequent over, regardless of the over of the innings.

>>>>

(ref:allFig) Distribution of energy consumption displayed as letter value plots, illustrating harmonies and clashes, and how mappings change emphasis: **a** weekday/weekend faceted by quarter-of-year produces a harmony, **b** quarter-of-year faceted by weekday/weekend produces a harmony, **c** month-of-year faceted by quarter-of-year produces a clash, as indicated by the empy sets and white space. Placement within a facet should be done for primary comparisons. For example, arrangement in **a** makes it easier to compare across weekday type (x-axis) within a quarter (facet). It can be seen that in quarter 2, there is more mass occupied the lower tail on the weekends (letter value E with depth 1/8) relative to that of the weekdays (letter value D 1/16), which corresponds to more days with lower energy use in this period.

@@

(ref:bothcust) Energy consumption of a single customer shown with different distribution displays, and granularity arrangements: hour of the day; and weekday/weekend. **a** The side-by-side boxplots make the comparison between day types easier, and suggest that there is generally lower energy use on the weekend. Interestingly, this is the opposite to what might be expected. Plots **b**, **c** examine the temporal trend of consumption over the course of a day, separately for the type of day. The area quantile emphasizes time, and indicates that median consumption shows prolonged high usage in the morning on weekdays. The violin plot emphasizes subtler distributional differences across hours: morning use on weekdays is bimodal.

@@

(ref:cricex) Examining distribution of runs per by innings, over of the innings and number of wickets in previous innings. Plot **a** displays distribution using letter value plots. A gradual upward trend in runs per over can be seen in innings 1, which is not present in innings 2. Plot **b** shows quantile plots of runs per over across an indicator of wickets in the previous over, faceted by current over. When a wicket occurred in the previous over, the runs per over tends to be lower throughout the innings.
